Experiment 2 is an exact replication of experiment 1, with one difference: the prime duration is set to 50ms instead of 33ms. The predictions are straightforward: if the FAE scales with prime duration, then the FAE in experiment 2 should be larger than the one observed in experiment 1. This result would explain why the FAE observed in experiment 1 was smaller than the ones observed in prior studies.

In order to facilitate the interpretation of the results of experiment 2, we calculate the 95\% *prediction intervals* [PI\; cf. @Patil2016; @SpenceStanley2016] for a replication of experiment 1 with the effective sample size of experiment 2 (cf. @sec-exp2-analysis-RT). The logic of the *prediction interval* is to establish a range of possible effect sizes that can be expected in a replication due only to sampling error. In order words, if the only difference between the original study and the replication study is sampling variability, we should expect the effect sizes of the replication study to fall within the 95\% PI set by the original study.

According to @tbl-pis, we can make the following predictions for experiment 2: (1) Non-word repetition effects in masked priming are thought to be independent of prime duration (mostly because they are at best negligible to begin with). Thus, as far as the non-words are concerned, experiment 2 can be seen as a simple replication of experiment 1. The $PI_{95\%}$ for non-words from experiment 1 is [-5ms 1ms]. Thus, we expect that the effect size for non-word repetition in experiment 2 could be as low a -5ms or as high as 1ms; any result beyond these bounds would strongly imply that the differences between the effect sizes of the two experiments cannot be explained simply in terms of sampling error. (2) Masked repetition effects for words, on the other hand, are thought to linearly increase in magnitude as a function of prime duration. Thus, we expect that in experiment 2 the repetition effect size for the LF and HF conditions should be superior to the upper bounds of the $PI_{95\%}$ of experiment 1. The $PI_{95\%}$ from experiment 1 for the the repetition of HF words is [15 21] ms; the one for the LF condition is [25 31] ms. Thus, if our prime duration manipulation is successful in experiment 2, we should expect the repetition effect for the HF condition to be larger than 21ms while the same effect for the LF condition should be larger than 31 ms. (3) It is presently unclear whether the FAE interacts with prime duration. Given the FAE $PI_{95\%}$ [6 14] ms from experiment 1, we can interpret the results of experiment 2 in the following manner: if the FAE in experiment 2 falls within the [6 14] ms interval, we have no particular reason to think that FAE interacts with prime duration, as this result is exactly what would have been predicted if experiment 2 was a direct replication of experiment 1. However, if the FAE in experiment 2 is larger than 14 ms (as the previous literature implies), then we have good reason to believe that the FAE linearly scales with prime duration, much like masked repetition priming effects do.

```{r}
#| echo: false
#| warning: false
#| error: false
#| label: tbl-pis
#| tbl-cap: "Prediction intervals calculated on the ANOVA-based means and standard deviations (SD) of the conditions tested in experiment 1."
#| tbl-pos: 'h'

pis %>% 
  mutate(across(c(6:7), round),
         pi.lb = paste0("[", pi.lb),
         pi.ub = paste0(pi.ub, "]")) %>%
  unite("95% PI", pi.lb:pi.ub, sep=" ") %>%
  gt() %>%
  cols_label(
    M = "mean",
    N = "experiment 1",
    rep.n = "experiment 2") %>%
  tab_spanner(
    label = "sample size",
    columns = c(4:5)
  )
  
```

## Methods {#sec-exp2-methods}

### Participants {#sec-exp2-methods-participants}

Two thousand and six hundred participants (1,551 females; _mean age_ = 39, _sd age_ = 12) were recruited on Prolific (<https://www.prolific.com>) with the same criteria specified for experiment 1 (@sec-exp1-methods-participants). Participants from experiment 1 were not allowed to enroll in this experiment. 

### Design {#sec-exp2-methods-design}

The experimental design was identical to experiment 1.

### Materials {#sec-exp2-methods-materials}

The experimental items were the same as experiment 1.

### Procedure {#sec-exp2-methods-proc}

Experiment 2 followed the same procedures as experiment 1 (see @sec-exp1-methods-proc). The only difference was the prime duration, which was set to 50 ms. The median time to finish the experiment was about 6 minutes, like in experiment 1.

## Data analysis {#sec-exp2-analysis}

The analysis and data pre-processing plan were the same as in experiment 1 (@sec-exp1-analysis). Analysis scripts and an abridged version of the data collected can be found online (<https://osf.io/k3gpc>), and consisted of `r format(nrow(exp2_rawdata.sub), big.mark=",", scientific=F)` observations in total.

### Step 1: subject and item performance {#sec-exp2-analysis-performance}

No item had an error rate superior to 12%, so no item was excluded from analysis. Thirty-nine subjects were removed because their error rate was above 30%. Thus, a total of `r format(nrow(exp2_data_step1), big.mark=",", scientific=F)` observations and `r format(exp2_step1_subj_remain, big.mark=',', scientific=F)` participants remained available for further analyses.

### Step 2: prime durations {#sec-exp2-analysis-primeTime}

Prime fluctuations were dealt with in the same way as in experiment 1 (@sec-exp1-analysis-primeTime). The mean (`r round(exp2_summary.primeTime$meanPrimeTime, 2)` ms, SD = `r round(exp2_summary.primeTime$sdPrimeTime, 2)` ms) and the median (`r median(exp2_rawdata.sub$primeTime)` ms) prime durations were close to the intended value (50 ms). Any trial whose prime duration was out of the 25-60ms range was excluded, resulting in the removal of `r round((exp2_primeTimeRangeSummary$n[1] + exp2_primeTimeRangeSummary$n[2])*100/nrow(exp2_rawdata.sub))`% of the trials. As compared to experiment 1, experiment 2 had a 7% larger percentage of removed out-of-range trials, the majority of which (`r round(exp2_primeTimeRangeSummary$range.percent[1])`%) were above the likely subliminal threshold of 60 ms. This is in contrast with the distribution of the prime durations in experiment 1, where the trials above the range amounted to `r round(exp1_primeTimeRangeSummary$range.percent[1], 2)`% of the dataset. As observed in previous studies and pilots conducted in our lab, this distribution suggests that setting the prime duration closer to either limit of a given range has the side effect of allowing for more fluctuations beyond either limit, thus potentially leading to greater data loss. No participant was excluded, and a total of `r format(exp2_step2_trials_remain, big.mark=",", scientific=F)` observations remained available for analysis.

### Step 3: RT analysis {#sec-exp2-analysis-RT}

After removing the incorrect responses, `r 100-round(exp2_step3_trials_remain*100/exp2_step2_trials_remain, 2)`% of the trials were excluded because their corresponding RT was below 200 ms or above 1800 ms. Finally, `r length(exp2_subj_filter_2)` subjects were removed because the number of trials within the same condition was less than 7. A total of `r format(exp2_final_trials_remain, big.mark=",", scientific=F)` observations and `r format(exp2_final_subj_remain, big.mark=",", scientific=F)` subjects remained available for statistical analysis.

## Results {#sec-exp2-results}

In the word analysis, the 2x2 repeated-measures ANOVA model revealed significant main effects (factor: *F*(1, 1923)=987.4, *p*<.0001; primetype: *F*(1, 1923)=1447, *p*<.0001) and interaction (*F*(1, 1923)=36.82, *p*<.0001). Planned comparisons confirmed statistically significant repetition priming effects for both word conditions ($MOP_{HF}$=26 ms, $CI_{95\%}$=[24 28], *t*(1923)=24.6, *p*<.0001; $MOP_{LF}$ = 35 ms, $CI_{95\%}$=[33 37], *t*(1923)=30.2, *p*<.0001), with the low-frequency word repetition priming effect being 10 ms larger than the high-frequency word repetition priming effect. This FAE effect was statistically significant ($M_{FAE}$=9 ms, $CI_{95\%}$=[6 12]), *t*(1920)=6.07, *p*<.0001). A very small but statistically significant inhibitory priming effect was observed in the non-word condition ($MOP_{NW}$=-4 ms, $CI_{95\%}$=[-6 -2], *t*(1923)=-3.53, *p*=.0004). The 2x2 GLMM analysis confirmed significant effects for condition ($\chi^2(1)=834.43, p<.0001$), primetype ($\chi^2(1)=3274.43, p<.0001$), and their interaction ($\chi^2(1)=219.76, p<.0001$). Similarly to experiment 1, the estimates of both models were comparable to one another (see @tbl-exp2-statsResults-anova and @tbl-exp2-statsResults-glmm)

In the error analysis, the 2x2 ANOVA showed significant effects for both main effects (factor: *F*(1, 2340)=392.5, *p*<.0001; primetype: *F*(1, 2340)=380.5, *p*<.0001) and interaction (*F*(1, 2340)=55.47, *p*<.0001). Planned comparisons confirmed significant priming effects in the form of fewer errors in repeated compared to unrelated trials in the word conditions (high: $M = 1.29, CI_{95\%}=[1.02 ~ 1.56], t(1923)=9.30, p<.0001$; low: $M = 3.05, CI_{95\%}=[2.67 ~ 3.43], t(1923)=15.8, p<.0001$), and an inverse effect in the non-word condition ($M = -0.35, CI_{}=[-0.64 ~ -0.054], t(1923)=-2.32, p=.002$). Similar results were obtained in the 2x2 GLMM error analysis (condition: $\chi^2(1)=4.95, p=.02$; primetype: $\chi^2(1)=101.53, p<.0001$; interaction: $\chi^2(1)=7.18, p=.007$). The GLMM estimated marginal means from the corresponding 3x2 GLMM model indicated the same trends as the ANOVA planned comparisons: (high: $\beta = 0.64, SE = 0.065, z=9.87, p<.0001$; low: $\beta = 0.87, SE = 0.051, z=17.14, p<.0001$), and an inhibitory effect in the non-word condition ($\beta = -0.08, SE = 0.033, z=-2.96, p=.02$).

\blandscape
```{r}
#| echo: false
#| message: false
#| error: false
#| warning: false
#| label: tbl-exp2-statsResults-anova
#| tbl-cap: "Experiment 2. Summary of the word priming results within the ANOVA model. *Legend.* MOP: magnitude of priming."
#| tbl-pos: 'h'

exp2_summary.results %>%
  mutate(t = round(t, 2)) %>%
  relocate(c("sd_unrelated", "mean.error_unrelated"), .before=gd.mean_related) %>%
  gt() %>%
  cols_label(
    CI = "95% CI",
    contains("mean") ~ "mean",
    contains("sd") ~ "SD", 
    contains("error") ~ "Error (%)"
  ) %>%
  tab_spanner(
    label = "unrelated RT",
    columns = c(2:4)
  ) %>%
  tab_spanner(
    label = "repetition RT",
    columns = c(5:7)
  ) %>%
  tab_spanner(
    label = 'priming effects',
    columns = c(9:12)
  ) %>%
  tab_spanner(
    label = md("_t_-test"),
    columns = c(13:15)
  ) %>%
  cols_label(
    sd = md("SD~p~")
  ) %>%
  cols_label(
    t = md("_t_"),
    p = md("_p_"),
  ) %>%
   sub_missing(
    missing_text = " "
  )
```

```{r}
#| echo: false
#| label: tbl-exp2-statsResults-glmm
#| tbl-cap: "Experiment 2. Summary of the word priming results based on the 2x3 GLMM model, with the same fixed and random effects as the 2x2 GLMM model, but including the non-word condition. *Legend.* MOP: magnitude of priming."
#| tbl-pos: 'h'

exp2_summary.results_glmm <- bind_rows(exp2_emmeans_glmm, exp2_emmeans_glmm_fae)
  
exp2_summary.results_glmm %>%
  gt() %>%
  cols_label(
    contains("CI") ~ "95% CI",
    contains("emmean") ~ "mean",
    contains("SE") ~ "SE"
  ) %>%
  tab_spanner(
    label = "unrelated RT",
    columns = c(2:4)
  ) %>%
  tab_spanner(
    label = "repetition RT",
    columns = c(5:7)
  ) %>%
  tab_spanner(
    label = 'priming effects',
    columns = c(8:10)
  ) %>%
   sub_missing(
    missing_text = " "
  )

```
\elandscape

## Discussion {#sec-exp2-discussion}

Experiment 2 aimed to replicate the FAE observed in Experiment 1 using an extended prime duration of 50 ms, and to assess whether this longer prime would amplify the effect. To achieve this, the same stimuli and sample size from Experiment 1 were employed, with the sole modification being the increased prime duration.

As predicted, as compared to experiment 1, we found an increase in the masked repetition priming effects for both word conditions in experiment 2, by approximately 8 ms for high-frequency ($M_{exp1} = 18ms ~ [16 ~ 20], M_{exp2} = 26ms ~ [24 ~ 28]$) and 7 ms for low-frequency ($M_{exp1} = 28ms ~ [26 ~ 30], M_{exp2} = 35ms ~ [33 ~ 37]$) words. Specifically, the repetition priming effect for high-frequency words was 26 ms, and for low-frequency words, 35 ms, both exceeding the upper bounds of the 95\% prediction interval from Experiment 1 (20 ms and 30 ms, respectively; see @tbl-pis). This indicates that the manipulation of prime duration was effective. Notably, and consistent with our predictions, the repetition effect for the non-word condition was -2 ms, which fell within the 95\% prediction interval of Experiment 1 (i.e., [-5 1]). Furthermore, the larger priming response for the high- and low-frequency word conditions in Experiment 2 was driven entirely by faster recognition of the related word pairs. The RT estimates for the unrelated pairs, along with their standard deviations and error rates, were virtually identically to those observed in Experiment 1, despite being from a completely different set of participants. This finding further corroborates the long-standing observation that the magnitude of the masked priming response is influenced by prime duration [@ForsterEtal2003].

Crucially, we were also able to replicate the FAE in Experiment 2. However, the FAE in this experiment was 9 ms ($CI_{95\%} = [6 ~ 12]$), which is nearly identical to the FAE observed in Experiment 1 and falls within its 95\% prediction interval, i.e. [6 14] ms. This suggests that, unlike the repetition effect in masked priming, the FAE does not interact with prime duration. Therefore, the difference in prime duration between Experiment 1 and prior studies in the literature does not appear to account for the observed discrepancies in the magnitude of the FAEs.

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| warning: false -->
<!-- #| error: false -->
<!-- #| label: fig-prime-distributions -->
<!-- #| fig-cap: "Distribution of the prime durations across the two datasets." -->

<!-- exps_data_final <- bind_rows(exp1_data_final_with.errors, exp2_data_final_with.errors, .id='experiment') %>% -->
<!--   mutate(experiment = factor(experiment)) -->

<!-- exps_data_final %>% -->
<!--   ggplot(aes(x=primeTime, linetype=experiment),)+ -->
<!--   geom_histogram(aes(x=primeTime, fill=experiment),  -->
<!--                  binwidth=1, alpha=0.8, position = "nudge")+ -->
<!--   #geom_density(stat="count")+ -->
<!--   scale_fill_manual(values=c("grey20", "grey60")) + -->
<!--   scale_x_continuous(breaks=seq(exp1_info$prime_dur_lb, exp1_info$prime_dur_ub, 5), -->
<!--                      limits=c(exp1_info$prime_dur_lb, exp1_info$prime_dur_ub))+ -->
<!--   theme_bw() +  -->
<!--   theme(panel.grid.major = element_blank(), -->
<!--       panel.grid.minor = element_blank(),  -->
<!--       panel.border = element_blank(), -->
<!--       legend.position="bottom") + -->
<!--   labs(x = "prime duration", y = "number of trials") -->

<!-- exps_prime.durations_tTest <- t.test(primeTime ~ experiment, data=exps_data_final, var.equal=T) -->

<!-- ``` -->

