Experiment 1 was designed to elicit a masked repetition priming response to both high- and low-frequency words. The required sample size to ensure adequate statistical power (>80%) was determined through a power analysis simulation (details and code available as supplemental material at https://osf.io/r7d2q/). For a 10 ms FAE, a sample size of 1,250 participants was identified as necessary to maintain 80% statistical power across a wide range of plausible standard deviations and correlation structures between conditions. Given the relatively untested nature of web browser-based masked priming experiments, we opted for a larger sample size of 2,600 participants. This decision was made to increase the likelihood of retaining at least 1,250 participants after applying exclusion criteria, while also guaranteeing a narrow margin of error around the estimated effect size [@Maxwell2008].

The prime duration was set at 33 ms for three main reasons. First, prior studies by @Angele2023, @Cayado2023, and @PetrosinoEtal2023 have demonstrated that a 33 ms prime duration is sufficient to elicit reliable repetition priming effects in web browser-based experiments. Second, this brief prime duration ensures that the prime word is almost never consciously perceived by participants [@ForsterEtal2003; @Nievas2010], providing a more accurate measure of early, presumably automatic processes involved in word recognition. Lastly, previous research on the *Labvanced* platform [@PetrosinoEtal2023] has shown that timing inaccuracies and missed screen refreshes can lead to an increased number of trials with actual prime durations exceeding the subliminal threshold (typically considered to be around 60 ms) when the prime duration is set to 50 ms. If a significant number of trials exceed the subliminal threshold, it may prompt participants to adopt experiment-wide strategies, ultimately compromising the observed masked priming response [@Zimmerman2012].

## Methods {#sec-exp1-methods}

### Preregistration {#sec-exp1-prereg}

We preregistered the results of the power analysis, the goals, the design and analysis plan for this experiment prior to data collection. The preregistration is available online (<https://doi.org/10.17605/OSF.IO/3NFQP>).

### Participants {#sec-exp1-methods-participants}

Two thousand and six hundred participants (1,445 females; _mean age_ = 42, _sd age_ = 14) were recruited on Prolific (<https://www.prolific.com>). Several criteria were selected to ensure recruitment of native speakers of U.S. English: participants had to be born in the Unites States of America, speak English as their first and only language, and have no self-reported language-related disorder. We encouraged participants to avoid any sort of distraction throughout the experiment, and to close any program that may be running in the background. However, because the experiment was run online, participants could not be monitored during data collection. Finally, to further reduce variability across participants' devices, we restricted the experiment to be run on Google Chrome only, which at the time of this writing is the most used browser worldwide [@w3counterGlobalStats], and reportedly performs better than any other across operating systems [see @LukacsGaspar2023].

### Design {#sec-exp1-methods-design}

The masked priming procedure relied on a lexical decision task (LDT), in which a 2 (frequency: _high_ vs _low_) x 2 (prime type: _repetition_ vs _unrelated_) factorial design was used. Both factors were manipulated within-subjects. The dependent variables were lexical decision latency (RT, in milliseconds) and error rate (in percentages).

### Materials {#sec-exp1-methods-materials}

One-hundred and four five-letter words, half of low frequency (between 7 and 24 occurrences per million in the SUBTLEX$_{US}$ corpus) and half of high frequency (between 57 and 2,961 occurrences per million in the SUBTLEX$_{US}$) were sampled from the English Lexicon Project [@balota2007]. From each condition, 26 words were selected to be presented as targets and related primes (the *repetition* condition), and the remaining 26 were presented as unrelated primes (the *unrelated* condition). All word items were also controlled for orthographic neighborhood (i.e., Coltheart's _N_): $t \approx 0$. All words used were monomorphemic nouns, adjectives, or verbs, thus excluding particles, prepositions, and derived or inflected forms.


```{r}
#| echo: false
#| warning: false
#| error: false
#| label: tbl-words_exp1
#| tbl-cap: "Experiment 1. Descriptive statistics of the word items used. For both frequency databases, the word frequencies were converted to per-million count to ensure cross-comparison."

words %>%
  group_by(freq.bin) %>%
  mutate(Freq_HAL.Pm = (Freq_HAL * 10^6)/(131 * 10^6)) %>%
  summarise(N = n(),
            minFreq= min(Freq_HAL.Pm, na.rm = T), maxFreq=max(Freq_HAL.Pm, na.rm = T),
            meanFreq = mean(Freq_HAL.Pm, na.rm = T), sdFreq = sd(Freq_HAL.Pm, na.rm = T),
            minSUBFreq= min(SUBTLWF, na.rm = T), maxSUBFreq=max(SUBTLWF, na.rm = T),
            meanSUBFreq = mean(SUBTLWF, na.rm = T), sdSUBFreq = sd(SUBTLWF, na.rm = T),
            minN = min(Ortho_N, na.rm=T), maxN = max(Ortho_N, na.rm=T),
            meanN = mean(Ortho_N, na.rm=T), sdN = sd(Ortho_N, na.rm = T)
            #meanRT=mean(as.numeric(I_Mean_RT), na.rm = T), sdRT=sd(as.numeric(I_Mean_RT), na.rm = T),
            #meanLength = mean(Length, na.rm = T), sdLength=sd(Length, na.rm = T)
            ) %>% 
  mutate(across(c(3:10), round), across(11:14, round, 2))  %>% 
  #mutate_if(needs_mutated, ~round(.)) %>%
  gt() %>%
  cols_label(
    freq.bin = "frequency", 
    minFreq = "min", maxFreq = "max", meanFreq = "mean", sdFreq = "SD", 
    minSUBFreq = "min", maxSUBFreq = "max", meanSUBFreq = "mean", sdSUBFreq = "SD",
    minN = "min", maxN = "max", meanN = "mean", sdN = "SD") %>%
  tab_spanner(
    label=md("**HAL**"),
    columns=3:6
  ) %>%
  tab_spanner(
    label=md("**SUBTLEX~US~**"),
    columns=7:10
  ) %>%
  tab_spanner(
    label=md("**Orthographic _N_**"),
    columns=11:14
  )
```

One-hundred and four five-letter, phono-orthographically legal non-words were randomly selected from the English Lexicon Project database as well. Half of them (i.e., 52) were randomly selected to be presented as targets; the other half was instead used as unrelated non-word primes. None of the non-words contained any existing English morpheme. Both the words and non-words used in the experiments are reported in the appendix at the end of the paper (after the References section). In addition, all items had an error rate in the ELP smaller than 10%, to ensure that they would all be clearly distinguishable from words by participants.

### Procedure {#sec-exp1-methods-proc}

Each recruited participant was assigned one of two word lists, which differed only in the relatedness of the prime with respect to the target; otherwise, the two lists presented the same set of target words and nonwords (i.e., 104 pairs for each list). In one list, the three conditions (the high- and low-frequency word conditions, and the non-word condition) had half of the target items being preceded by themselves (the *repetition* condition) and half of the target items being preceded by one of the unrelated primes belonging to the same frequency bin (the *unrelated* condition). In the other list, these assignments were reversed. The order of stimulus presentation was randomized for each participant.

After being recruited in the _Prolific_ online platform, participants were asked to click on a link redirecting them to the Labvanced online service. During the experiment, they were asked to perform a lexical decision task by pressing either the 'J' (for word) or 'F' (for non-word) keys on their keyboard. Each trial consisted of three different stimuli appearing at the center of the screen: a series of five hashes (#####) presented for 500 ms, followed by a prime word presented for 33 ms, and finally the target word; the target word disappeared from the screen as soon as a decision was made.

Participants were given 5 breaks throughout the experiment. When the experiment was over, the participants were then redirected to Prolific in order to validate their submission. The median time to finish the experiment was 6 minutes and participants were paid a rate of 9 GBP/hour.

## Data analysis {#sec-exp1-analysis}

Our preregistered data analysis plan established that the data would be analyzed using a ***2x3*** repeated factors ANOVA (condition, 3 levels: high vs. low vs. non-word; primetype, 2 levels: unrelated vs. repetition), with only subjects as random factors, followed by planned comparisons between related and unrelated primetype for each condition. The choice of this statistical model follows the recommendations of @Raaijmakers1999 and @Raaijmakers2003, and allowed us to carry out the power analysis simulations that informed our sample size choice, since models that include items as random factors or even as crossed random factors are too computationally expensive to run in large scale statistical power simulations. However, since these models have become more widely adopted, we also present an analysis using Generalized Linear Mixed Model (GLMM) analysis on the raw RT and the raw accuracy data. Unlike linear mixed-effect models, GLMMs do not assume normal distribution of the data, and are therefore particularly useful for non-normally distributed data such as RTs and accuracy. For the RT data, a Gamma distribution was used; for the accuracy data, a binomial distribution was used instead. Both models used an identity link between fixed effects and the dependent variable [@LoAndrews2015]. To prevent converge failure, the model was kept as simple as possible, with *condition*, *primetype* and their interaction as fixed effects. Subjects and items were crossed as random effects. Before fitting the model, the contrasts were also set to sum-to-zero contrasts (i.e., by using the R function `contr.sum()`) to facilitate interpretation of main and interaction effects. The fitting was performed by using the `lme4` R-package [@BatesEtal2015] with the Laplace approximation technique, using 1 million iterations and the BOBYQA optimizer to help convergence. The function `Anova()` from the `car` R-package [@FoxWeisberg2019] was used to obtain estimates and probability values for fixed effects calculated for Type-III sums of squares. Estimated marginal means were then calculated with the function `emmeans()` from the the `emmeans` R-package [@Lenth2024].

Analysis scripts and an abridged version of the data collected can be found online (<https://osf.io/vn3r2>), and consisted of `r format(nrow(exp1_rawdata.sub), big.mark=",", scientific=F)` observations in total. The pre-processing of the data consisted of three separate steps.

### Step 1: subject and item performance {#sec-exp1-analysis-performance}

Item and subject error rates were calculated. The item error rate was never above 14%, so no item was excluded from analysis. `r exp1_info$n_recruited - exp1_step1_subj_remain` subjects were removed because their error rate was above 30%. Thus, a total of `r format(nrow(exp1_data_step1), big.mark=",", scientific=F)` observations and `r format(exp1_step1_subj_remain, big.mark=',', scientific=F)` participants were included in further analyses.

### Step 2: prime durations {#sec-exp1-analysis-primeTime}

During the experiment, the duration of presentation of the prime word was recorded for every trial. Both the mean (mean = `r round(exp1_summary.primeTime$meanPrimeTime)` ms; sd=`r round(exp1_summary.primeTime$sdPrimeTime, 2)` ms) and the median (`r median(exp1_rawdata.sub$primeTime)` ms) of the prime duration confirm some imprecision in the duration of the presentation of the prime during the experiment. This was expected and likely due to the inherent difficulty with timing precision of visual presentations in web browsers and the great variation of computer hardware and internet connections used by the participants. Both of these issues may be impossible to control, at least at the current state of browser development. However, in masked priming, in which the duration of the prime is an essential part of the design, such fluctuations may indeed hinder proper elicitation of the priming response. Thus, we only kept trials whose prime durations were within a pre-set range from the intended prime duration of 33 ms. Taking a standard 60-Hz monitor as reference, the lower and the upper bounds were set respectively at `r exp1_info$prime_dur_lb` ms (i.e., the intended prime duration minus half of a full refresh cycle: $33-8~ ms$; noting that @Angele2023 already showed that no repetition priming effects are obtained with a 16.7ms prime duration) and `r exp1_info$prime_dur_ub` ms (i.e., the commonly accepted upper threshold of subliminal processing), in an attempt to remove any trial that could have been consciously perceived by participants. This cut-off removed `r round((exp1_primeTimeRangeSummary$n[1] + exp1_primeTimeRangeSummary$n[2])*100/nrow(exp1_rawdata.sub))` % of the trials. No participant was excluded, for a total of `r format(exp1_step2_trials_remain, big.mark=",", scientific=F)` observations.

### Step 3: RT analysis {#sec-exp1-analysis-RT}

After removing the incorrect responses, `r 100-round(exp1_step3_trials_remain*100/exp1_step2_trials_remain, 2)`% of the trials were excluded if their corresponding RT was below 200 ms or above 1800 ms. Finally, `r length(exp1_subj_filter_2)` subjects were removed because the number of trials within the same condition was less than 7 (i.e., about half of the total number of trials being presented within the same condition, i.e. 13). A total of `r format(exp1_final_trials_remain, big.mark=",", scientific=F)` observations and `r format(exp1_final_subj_remain, big.mark=",", scientific=F)` subjects were included in the statistical analysis below. 

## Results {#sec-exp1-results}

Priming effects were calculated by subtracting the mean RT to the related condition to the mean RT from the unrelated condition. We ran two different analyses, for both RT and error data. 

The preregistered 2x3 repeated-measures ANOVA (condition, 3 levels: high vs. low vs. non-word; primetype, 2 levels: unrelated vs. repetition) on the by-subject averaged RT data revealed significant main effects (condition: *F*(1, 2340)=1572, *p*<.0001; primetype: *F*(1, 2340)=1113, *p*<.0001) and their interaction *F*(1, 2340)=52.48, *p*<.0001). Planned comparisons confirmed statistically significant repetition priming effects for both word conditions (*MOP_HF*=18 ms, *CI_95%*=[16 20], *t*(2340)=19.7, *p*<.0001; *MOP_LF* = 28 ms, *CI_95%*=[26 30], *t*(2340)=27.8, *p*<.0001), with the low-frequency word repetition priming effect being 10 ms larger than the high-frequency word repetition priming effect. This FAE effect was statistically significant (*M_FAE*=10 ms, *CI_95%*=[7 13], *t*(2340)=7.24, *p*<.0001). A very small, but statistically significant inhibitory priming effect was observed in the non-word condition (*MOP_NW*=-2 ms, *CI_95%*=[-4 0], *t*(2340)=-2.33, *p*<.0001). Similarly, the GLMM analysis confirmed that both main effects (condition: $\chi^2=2978.35, p<.0001$; primetype: $\chi^2=1531.12, p<.0001$), and their interaction ($\chi^2=870.27, p<.0001$) were significant. As reported in @tbl-exp1-statsResults-anova and @tbl-exp1-statsResults-glmm, the estimates based on the ANOVA and GLMM models overlap with one another, which suggest the two models to be substantially equivalent.

In the error analysis, the ANOVA revealed significant effects for both main effects (condition: *F*(1, 2340)=392.5, *p*<.0001; primetype: *F*(1, 2340)=380.5, *p*<.0001) and interaction *F*(1, 2340)=55.47, *p*<.0001). Planned comparisons revealed significant priming effects in the form of fewer errors in the repeated compared to unrelated trials in the all conditions (high: *t*(2340)=9.95, *p*<.0001; low: *t*(2340)=16.9, *p*<.0001; non-word: *t*(2340)=-3.27, *p*=.001). As in the ANOVA error analysis, both main effects (condition: $\chi^2=30.45, p<.0001$; primetype: $\chi^2=108.88, p<.0001$) and their interaction ($\chi^2=307.57, p<.0001$) were significant.

\blandscape
```{r}
#| echo: false
#| message: false
#| error: false
#| warning: false
#| label: tbl-exp1-statsResults-anova
#| tbl-cap: "Experiment 1. Summary of the word priming results calculated within the ANOVA model. *Legend.* MOP: magnitude of priming."
#| tbl-pos: 'h'

exp1_summary.results %>%
  relocate(c("sd_unrelated", "mean.error_unrelated"), .before=gd.mean_related) %>%
  gt() %>%
  cols_label(
    CI = "95% CI",
    contains("mean") ~ "mean",
    contains("sd") ~ "SD", 
    contains("error") ~ "Error (%)"
  ) %>%
  tab_spanner(
    label = "unrelated RT",
    columns = c(2:4)
  ) %>%
  tab_spanner(
    label = "repetition RT",
    columns = c(5:7)
  ) %>%
  tab_spanner(
    label = 'priming effects',
    columns = c(9:12)
  ) %>%
  tab_spanner(
    label = md("_t_-test"),
    columns = c(13:15)
  ) %>%
  cols_label(
    sd = md("SD~p~")
  ) %>%
  cols_label(
    t = md("_t_"),
    p = md("_p_"),
  ) %>%
   sub_missing(
    missing_text = " "
  )
```

```{r}
#| echo: false
#| label: tbl-exp1-statsResults-glmm
#| tbl-cap: "Experiment 1. Summary of the word priming results calculated within the GLMM model. *Legend.* MOP: magnitude of priming."
#| tbl-pos: 'h'

exp1_summary.results_glmm %>%
  gt() %>%
  cols_label(
    contains("CI") ~ "95% CI",
    contains("emmean") ~ "mean",
    contains("SE") ~ "SE"
  ) %>%
  tab_spanner(
    label = "unrelated RT",
    columns = c(2:4)
  ) %>%
  tab_spanner(
    label = "repetition RT",
    columns = c(5:7)
  ) %>%
  tab_spanner(
    label = 'priming effects',
    columns = c(8:10)
  ) %>%
   sub_missing(
    missing_text = " "
  )

```
\elandscape

## Discussion {#sec-exp1-discussion}

Experiment 1 was designed to investigate whether Frequency Attenuation Effects (FAE) can be detected in masked priming with a prime duration of 33 ms. Our findings replicated the results of @Angele2023, @Cayado2023, and @PetrosinoEtal2023, showing statistically significant masked repetition priming effects in a web browser-based experiment for both high- and low-frequency words. Crucially, the results also confirmed the presence of an FAE, evidenced by a statistically significant interaction: the priming effect for low-frequency words was 10 ms larger than that for high-frequency words. Additionally, due to the large sample size, the FAE was estimated with high precision, yielding a narrow margin of error of 3 ms. Interestingly, in addition to the reliable FAE, our results align with previous studies [@Cayado2023; @PetrosinoEtal2023] in showing a minimal inhibitory masked repetition priming effect for non-words, with very high precision. The 95% confidence interval (CI) suggests that the plausible range for the masked repetition priming effect for non-words, at a prime duration of 33 ms, is between -4 ms and 0 ms.

These findings help dispel doubts about whether an FAE can be observed in masked priming, challenging earlier studies that failed to detect this effect. However, the results also raise questions about the size and malleability of the FAE. The FAE observed in Experiment 1 (10 ms) is substantially smaller than those reported in other studies where significant FAEs were found [e.g., @Kinoshita2006], where the effect was around 30 ms. What could account for this discrepancy in effect size? One possibility is that the 30 ms estimate is inflated due to the lower statistical power in many previous experiments. As noted in the introduction, low power can lead to overestimated effect sizes when results are statistically significant. When examining the range of effect sizes in @tbl-litReview, we find that the average FAE is approximately 15 ms, which is half the size of those observed when statistically significant results were obtained. This 15 ms estimate is closer to the 10 ms effect size observed in Experiment 1, but still somewhat larger. Thus, there appears to be a potential discrepancy between the FAE effect sizes reported in the literature and the estimate from Experiment 1.

One possible explanation for this discrepancy is the difference in prime duration used in Experiment 1 (33 ms) compared to most previous studies ($\geq 50 ms$). It is plausible that the FAE scales linearly with prime duration in masked priming, similar to how repetition priming effects increase with longer prime durations. In fact, this is an actual prediction of *IA* models [see @GraingerEtal2012]. Fortunately, @Angele2023 and @PetrosinoEtal2023 have already demonstrated that web browser-based experiments can replicate the proportionality between prime duration and the magnitude of masked identity priming, consistent with @ForsterEtal2003. Both studies showed that increasing the prime duration from 33 ms to 50 ms resulted in larger identity priming effects. This provides an opportunity to test whether the discrepancy in FAE effect size between Experiment 1 and prior literature is due to the shorter prime duration used in our study. The goal of Experiment 2, therefore, is to leverage the established linear scaling of repetition effects with prime duration in masked priming to determine whether extending the prime duration also increases the magnitude of the FAE.
